---
  title: "penguin_analysis"
author: "Catherine Stephenson"
date: "2024-10-16"
output: html_document
---
  ```{r}r
library(tidyverse)
library(palmerpenguins)
library(here)
library(janitor)
```

do not install packages everytime as the package may have changed

use library for loading in packages the same

use here() instead of setwd - privacy issue and more secure
```{r}
here()
```
load libraries

```{r}
write.csv(penguins_raw, here("PenguinData","penguins_raw.csv"))
```

never edit a file in excel makes it unreproducible

never override data by saving it as the same thing (if accidentally done then reload the read only saved copy)

cleaning the data table by removing columns in the penguin raw data:
  
  
  
  ```{r}
colnames(penguins_raw)
penguins_clean <- select(penguins_raw, -Comments)
colnames(penguins_clean)
```
same for two columns at once (how to not overwrite):
  ```{r}
penguins_cleaner <- penguins_raw %>%
  clean_names
select(-Comments) %>% #using piping to remove columns %>% means and then
  select(-starts_with("Delta"))



```
```{r}
#we can make a new function called cleaning_penguin_columns

#head(penguins_clean)

library(tidyverse)
library(palmerpenguins)
library(here)
library(janitor)

here()

write.csv(penguins_raw, here("data","penguins_raw.csv"))


cleaning_penguin_columns <-function(raw_data){
  raw_data %>%
    clean_names()%>%
    select(-comments)%>%
    select(-starts_with("delta"))
  
}

file.create(here("functions", "cleaning.R"))
```

to load the function file:
  
  

```{r}


source(here("functions", "cleaning.R"))

penguins_raw <- read.csv(here("data","penguins_raw.csv"))

cleaning_penguin_columns <-function(raw_data){
  print("Cleaned names, removed comments, removed empty rows and cols, removed delta")
  raw_data %<%
    clean_names() %>%
    select
  shorten_species() %>%
    remove_empty_columns_rows()
}
```


  
  we need to install renv
  
  ```{r}
  install.packages("renv")
  
  
  ```
  
  ```{r}
  renv::diagnostics
  
  ```
  ```{r}
  
  install.packages("table1")
  renv::snapshot
  ```
  
Now to part 2:reproducible figures

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)

source(here("functions", "cleaning.R"))

```

creating exploratory figures

- shows raw data ad is a good first step before modelling

```{r}
penguins_raw <- read_csv(here("data", "penguins_raw.csv"))

penguins_clean <- penguins_raw %>%
  clean_column_names() %>%
  remove_columns(c("comments", "delta")) %>%
  shorten_species() %>%
  remove_empty_columns_rows()

names(penguins_clean)
```
```{r}
flipper_boxplot <- ggplot(
  data = penguins_clean, 
  aes(x = species, 
      y = flipper_length_mm)) +
  geom_boxplot()

#to remove NAs:

penguins_flippers <- penguins_clean %>%
  select(species, flipper_length_mm) %>%
  drop_na()

flipper_boxplot
```

to make this look better:

```{r}
flipper_boxplot <- ggplot(
    data = penguins_flippers, 
    aes(x = species, 
        y = flipper_length_mm)) +
    geom_boxplot(aes(color = species), 
                show.legend = FALSE) +
    geom_jitter(aes(color = species), 
                show.legend = FALSE)
                          

flipper_boxplot
```
  
  however randomness is not reproducible (doesnt matter here, however)
  
  use a random_seed to create same randomness:
  
```{r}
flipper_boxplot <- ggplot(
    data = penguins_flippers, 
    aes(
        x = species, 
        y = flipper_length_mm)) +
    geom_boxplot(
        aes(color = species), 
        width = 0.3, 
        show.legend = FALSE) +
    geom_jitter(
        aes(color = species),
        show.legend = FALSE, 
        position = position_jitter(
            width = 0.2, 
            seed = 0))

flipper_boxplot
     
```
  
  to alter the transparency of elements - change the alpha value - 1= opaque, 0= transparent
  
```{r}
flipper_boxplot <- ggplot(
    data = penguins_flippers, 
    aes(
        x = species, 
        y = flipper_length_mm)) +
    geom_boxplot(
        aes(color = species), 
        width = 0.3, 
        show.legend = FALSE) +
    geom_jitter(
        aes(color = species),
        alpha = 0.3,
        show.legend = FALSE, 
        position = position_jitter(width = 0.2, seed = 0))

flipper_boxplot
     
```
  fixing the labels:
  
```{r}
flipper_boxplot <- ggplot(
    data = penguins_flippers, 
    aes(x = species, 
        y = flipper_length_mm)) +
    geom_boxplot(aes(color = species), 
                width = 0.3, 
                show.legend = FALSE) +
    geom_jitter(aes(color = species), 
                alpha = 0.3,
                show.legend = FALSE,
                position = position_jitter(
                    width = 0.2, 
                    seed = 0)) +
    labs(x = "Species",
         y = "Flipper Length (mm)")

flipper_boxplot
     
```
  changing the colour scheme:
  
```{r}
species_colours <- c("Adelie" = "darkorange", 
                    "Chinstrap" = "purple", 
                    "Gentoo" = "cyan4")

flipper_boxplot <- ggplot(
    data = penguins_flippers, 
    aes(
        x = species, 
        y = flipper_length_mm)) +
    geom_boxplot(
        aes(color = species), 
            width = 0.3, 
            show.legend = FALSE) +
    geom_jitter(
        aes(color = species), 
        alpha = 0.3,
        show.legend = FALSE,
        position = position_jitter(
            width = 0.2, 
            seed = 0)) +
    scale_color_manual(values = species_colours) +
    labs(x = "Species",
                               y = "Flipper Length (mm)")

flipper_boxplot
```
 we can change the theme to remove the background:
 
```{r}
species_colours <- c("Adelie" = "darkorange", 
                    "Chinstrap" = "purple", 
                    "Gentoo" = "cyan4")

flipper_boxplot <- ggplot(
    data = penguins_flippers, 
    aes(
        x = species, 
        y = flipper_length_mm)) +
    geom_boxplot(
        aes(color = species), 
        width = 0.3, 
        show.legend = FALSE) +
    geom_jitter(
        aes(color = species), 
        width = 0.3, 
        alpha = 0.3,
        show.legend = FALSE) +
    scale_color_manual(
        values = species_colours) + 
    labs(
        x = "Species",
        y = "Flipper Length (mm)") + 
    theme_bw()

flipper_boxplot
```
 this code can be turned into a function:
 
 - see the plotting.R script
 - have produced a function that is flexible for other plotting
 
```{r}
plot_boxplot(penguins_clean, species, body_mass_g, "Species", "Body Mass (g)", species_colours)

     
```
 can now use this function to make 4 subplots
 
```{r}
# 2 by 2 grid subplots
# options(repr.plot.width = 8, repr.plot.height = 8, repr.plot.res = 300)

library(patchwork)

# Create individual subplots with different y-axis columns
plot1 <- plot_boxplot(
    penguins_clean, species, flipper_length_mm, 
    "Penguin Species", "Flipper Length (mm)", 
    species_colours)

plot2 <- plot_boxplot(
    penguins_clean, species, culmen_length_mm, 
    "Penguin Species", "Bill Length (mm)", 
    species_colours)

plot3 <- plot_boxplot(
    penguins_clean, species, culmen_depth_mm, 
    "Penguin Species", "Bill Depth (mm)", 
    species_colours)

plot4 <- plot_boxplot(
    penguins_clean, species, body_mass_g, 
    "Penguin Species", "Body Mass (g)", 
    species_colours)

# Combine plots in a 2x2 grid
combined_plot <- (plot1 | plot2) / (plot3 | plot4)
combined_plot
```

saving plots into the figures file as pdf:

```{r}
library(ragg)
library(svglite)

flipper_boxplot <- plot_boxplot(penguins_clean, 
                                species, flipper_length_mm, 
                                "Species", "Flipper Length (mm)", 
                                species_colours)

agg_png("figures/penguin_boxplot_default.png", 
        width = 20, 
        height = 20, 
        units = "cm", 
        res = 300, 
        scaling = 2) #changing the scaling or resolution by changing this etc
print(flipper_boxplot)
dev.off()
```
 this means that our posters and report figures are identical but different sizes if we change these units
 
 to make a powerpoint figure:
 
```{r}
flipper_boxplot <- plot_boxplot(penguins_clean, 
                                species, flipper_length_mm, 
                                "Species", "Flipper Length (mm)", 
                                species_colours)

agg_png("figures/penguin_boxplot_powerpoint.png", 
        width = 20, 
        height = 20, 
        units = "cm", 
        res = 300, 
        scaling = 3)
print(flipper_boxplot)
dev.off()
     
```
 
 we can also produce vector graphs instead of PNG for higher resolution
 - always better to use vectors instead of PNG
 
```{r}
flipper_boxplot <- plot_boxplot(penguins_clean, 
                                species, flipper_length_mm, 
                                "Species", "Flipper Length (mm)", 
                                species_colours)
inches_conversion = 2.54
svglite("figures/penguin_boxplot_vector.svg", 
        width = 20 / inches_conversion, 
        height = 20 / inches_conversion, 
        scaling = 2)
print(flipper_boxplot)
dev.off()
```
 we can also make a function for saving our plots (in the plotting script)
 
```{r}
save_flipper_plot_png(flipper_boxplot, 
                       here("figures", "flipper_boxplot_report.png"), 
                       size = 20, res = 300, scaling = 2)

save_flipper_plot_png(flipper_boxplot, 
                       here("figures", "flipper_boxplot_poster.png"), 
                       size = 40, res = 300, scaling = 4)

save_flipper_plot_png(flipper_boxplot, 
                       here("figures", "flipper_boxplot_powerpoint.png"), 
                       size = 20, res = 300, scaling = 3)



```
```{r}
save_flipper_plot_svg(flipper_boxplot, 
                       here("figures", "flipper_boxplot_report.svg"), 
                       size = 20, scaling = 2)

save_flipper_plot_svg(flipper_boxplot, 
                       here("figures", "flipper_boxplot_poster.svg"), 
                       size = 40, scaling = 4)

save_flipper_plot_svg(flipper_boxplot, 
                       here("figures", "flipper_boxplot_powerpoint.svg"), 
                       size = 20, scaling = 3)


     
```
 
  